CrypTool Script \\ Mathematics and Cryptography

### 2002-07-01 script/en/chapters/numbertheory.tex
appendix 3.d english; typo in german 3

Cryptographically not only the uniqueness of the inverse is important, but also that the
set of possible values is exhausted.
\begin{theorem}\label{thm-zth-exhperm}
For $a,i\in \{1, \cdots, m-1 \}$ with ${\rm~gcd} (a,m)=1),$ then takes the product $a*i {\rm ~mod ~} m $ for a 
certain number $a$ all values from $\{1, \cdots ,m-1 \} $ (exhaustive permutation \index{Permutation} of the Length $m-1$)\footnote{See
also theorem \ref{thm-zth-ordp} in \hyperlink{Kap_3_9}{chapter 3.9 Multiplicative order and primitive roots}.}.
\end{theorem}
\vskip +10 pt
The following three examples\footnote{Mathematica can be used to calculated the tables for these examples as follows: \\
{\tt m = 17; iWidth = 18; iFactor1 = 5; iFactor2 = 6; \\
Print[ ``i'', Table[ i, \{i, 1, iWidth \} ] ]; \\
Print[ iFactor1, ``*i'', Table[ iFactor1*i, \{i, 1, iWidth \} ] ]; \\
Print[ ``remainder'', Table[ Mod[iFactor1*i, m], \{i, 1, iWidth \} ] ]; \\
Print[ iFactor2, ``*i'', Table[ iFactor2*i, \{i, 1, iWidth \} ] ];  \\
Print[ ``remainder'', Table[ Mod[iFactor2*i, m], \{i, 1, iWidth \} ] ]; 
} } illustrate the properties of multiplicative inverses.

@@

{
\begin{table}[h]
\subsubsection*{Multiplication table modulo $17$ (for $a=5$ and $a=6$)}
\begin{center}
\begin{tabular}{|l||@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}||@{\:}c@{\:}|@{\:}c@{\:}|}
\hline 
i                   & 1  & 2  & 3  & 4  & 5  & 6  & 7  & 8  & 9 & 10 & 11 & 12 & 13 & 14 & 15 & 16  & 17 & 18 \\
\hline
\hline  
$5*i$                & 5 & 10 & 15 & 20 & 25 & 30 & 35 & 40 & 45 & 50 & 55 & 60 & 65 & 70 & 75 & 80  & 85 & 90   \\
remainder                & 5 & 10 & 15  & 3  & 8 & 13  & {\bf 1}  & 6 & 11 & 16  & 4  & 9 & 14  & 2  & 7 & 12 & 0  & 5   \\
\hline
$6*i$                 & 6 & 12 & 18 & 24 & 30 & 36 & 42 & 48 & 54 & 60 & 66 & 72 & 78 & 84 & 90 & 96 & 102 & 108   \\
remainder                & 6 & 12  & {\bf 1}  & 7 & 13  & 2  & 8 & 14  & 3  & 9 & 15  & 4 & 10 & 16  & 5 & 11 & 0  & 6   \\
\hline
\end{tabular}
\end{center} 
\end{table}
}
Between $i=1, \cdots, m$, all values between $0, \cdots, m-1$ occur for the remainders, because both $5$ and $6$ are 
also relatively prime\index{Relatively prime} to the modulus $m=17$.
\enlargethispage{0.5cm}

@@

\begin{table}[h]
\subsubsection*{Multiplication table modulo $13$ (for $a=5$ and $a=6$)}
\begin{center}                                                                          
\begin{tabular}{|l||@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}||@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|}
\hline 
i                    & 1  & 2  & 3  & 4  & 5  & 6  & 7  & 8  & 9 & 10 & 11 & 12 & 13 & 14 & 15 & 16  & 17  & 18 \\
\hline 
\hline 
$5*i$                  & 5 & 10 & 15 & 20 & 25 & 30 & 35 & 40 & 45 & 50 & 55 & 60 & 65 & 70 & 75 & 80 & 85  & 90 \\
remainder                 & 5 & 10  & 2  & 7  & 12  & 4 & 9  & {\bf 1}  & 6  & 11 & 3  & 8  & 0 & 5  & 10  & 2   & 7   & 12 \\
\hline 
$6*i$                  & 6 & 12 & 18 & 24 & 30 & 36 & 42 & 48 & 54 & 60 & 66 & 72 & 78 & 84 & 90 & 96 & 102 & 108 \\
remainder                 & 6  & 12  & 5  & 11  & 4  & 10  & 3  & 9  & 2  & 8  & {\bf 1}  & 7  & 0  & 6  & 12  & 5   & 11   & 4 \\
\hline 
\end{tabular}
\end{center} 
\end{table}
}

@@

and the squares (mod $103$) can be calculated beforehand
\begin{eqnarray*}
87^2 & \equiv & 50 {\rm ~(mod~}103),\\
87^4 & \equiv & 50^2 \equiv 28 {\rm ~(mod~}103), \\
87^8 & \equiv & 28^2 \equiv 63 {\rm ~(mod~}103), \\
87^{16} & \equiv & 63^2 \equiv 55 {\rm ~(mod~}103),\\
87^{32} & \equiv & 55^2 \equiv 38 {\rm ~(mod~}103).
\end{eqnarray*}
we have\footnote{Using Mathematica this is calculated in less than 
a second on a Pentium 3: \\
{\tt Mod[{${\rm 87}^\wedge$43, ${\rm 87}^\wedge$2, ${\rm 87}^\wedge$4, ${\rm 87}^\wedge$8, ${\rm 87}^\wedge$16, ${\rm 87}^\wedge$32}, 103]
 = \{85, 50, 28, 63, 55, 38\}.
} % \tt
}:
\begin{eqnarray}
87^{43} & \equiv & 87^{32+8+2+1}{\rm ~(mod~}103) \nonumber \\
        & \equiv & 87^{32} * 87^8 * 87^2 * 87 {\rm ~(mod~}103) \nonumber \\ 
    & \equiv & 38 * 63 * 50 * 87 \equiv 85 {\rm ~(mod~}103). \nonumber
\end{eqnarray}

@@

{\bf Example 1:} \\
The following table\footnote{Mathematica can be used to calculate this as follows: \\
{\tt m=11;  Table[ Mod[${\rm a}^\wedge$i, m], \{a, 1, m-1\}, {i, 1, m-1} ].} } shows the values $a^i$ mod $11$ for the exponents $i = 1, 2, \cdots, 10$ and for the bases 
$a =  1, 2, \cdots, 10$ as well as the resulting value $ord_{11}(a)$ for each $a$:

{
\begin{table}[h]
\subsubsection*{Values of $a^i {\rm ~mod~} 11,  1 \leq a,i<11$ and according order of $a$ mod $m$:}
\begin{center}                                                                          
\begin{tabular}{|l||c|c|c|c|c|c|c|c|c|c|c|c|c|c|}
\hline
              & i=1 & i=2 & i=3 & i=4 & i=5 & i=6 & i=7 & i=8 & i=9 & i=10  & $ord_{11}(a)$\\
\hline
\hline
$a=1$           & 1  & 1    & 1  & 1    & 1    & 1    & 1  & 1    & 1  & 1     & 1   \\
\hline
$a=2$           & 2  & 4    & 8  & 5   & 10    & 9    & 7  & 3    & 6  & 1    & 10  \\
\hline
$a=3$           & 3  & 9    & 5  & 4 & {\bf 1} & 3    & 9  & 5    & 4  & 1     & 5   \\
\hline
$a=4$           & 4  & 5    & 9  & 3 & {\bf 1} & 4    & 5  & 9    & 3  & 1    & 5 \\
\hline
$a=5$           & 5  & 3    & 4  & 9 & {\bf 1} & 5    & 3  & 4    & 9  & 1    & 5   \\
\hline
$a=6$           & 6  & 3    & 7  & 9   & 10    & 5    & 8  & 4    & 2  & 1    & 10  \\
\hline
$a=7$           & 7  & 5    & 2  & 3   & 10    & 4    & 6  & 9    & 8  & 1    & 10  \\
\hline
$a=8$           & 8  & 9    & 6  & 4   & 10    & 3    & 2  & 5    & 7  & 1    & 10  \\
\hline
$a=9$           & 9  & 4    & 3  & 5 & {\bf 1} & 9    & 4  & 3    & 5  & 1    & 5   \\
\hline
$a=10$         & 10  & 1   & 10  & 1   & 10    & 1   & 10  & 1   & 10  & 1    & 2   \\
\hline
\end{tabular}
\end{center}
\end{table}
}

@@

{\bf Example 2:} \\
The following table\footnote{Mathematica can be used to calculate this as follows (no Do loop can be used in Print and each
Print command outputs a new line at the end):\\
{\tt 
m = 45; \\
Do[ Print[ Table[ Mod[${\rm a}^\wedge$i, m], \{i, 1, 12\} ], \\
'', '', MultiplicativeOrder[a, m, 1], \\
'', '', EulerPhi[m] ], \\
\{a, 1, 12\} \\
]; } % \tt 
} shows the values $a^i$ mod $45$ for the exponents $i = 1, 2, \cdots, 12$ and for the bases 
$a =  1, 2, \cdots, 12$ as well as the resulting value $ord_{45}(a)$ for each $a$:

{ % \small
\begin{table}[h]
\subsubsection*{Values of $a^i{\rm ~mod~}45, 1\leq a,i<13$:}
\begin{center}                                                                          
\begin{tabular}{|l||c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}
\hline
 $a\setminus i$ & 1            & 2            & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12     & $ord_{45}(a)$       & $J(45)$ \\
\hline
\hline                                                       
1             & 1              & 1   & 1   & 1   & 1   & 1   & 1   & 1   & 1    & 1    & 1    & 1 & 1              & 24  \\
\hline
2             & 2              & 4   & 8  & 16  & 32  & 19  & 38  & 31  & 17   & 34   & 23    & 1 & 12             & 24 \\
\hline
3             & 3              & 9  & 27  & 36  & 18   & 9  & 27  & 36  & 18    & 9   & 27   & 36  & ---            & 24 \\
\hline
4             & 4             & 16  & 19  & 31  & 34   & 1   & 4  & 16  & 19   & 31   & 34    & 1  & 6              & 24 \\
\hline
5             & 5             & 25  & 35  & 40  & 20  & 10   & 5  & 25  & 35   & 40   & 20   & 10  & ---            & 24 \\
\hline
6             & 6             & 36  & 36  & 36  & 36  & 36  & 36  & 36  & 36   & 36   & 36   & 36  & ---            & 24 \\
\hline
7             & 7              & 4  & 28  & 16  & 22  & 19  & 43  & 31  & 37   & 34   & 13    & 1  & 12             & 24 \\
\hline
8             & 8             & 19  & 17   & 1   & 8  & 19  & 17   & 1   & 8   & 19   & 17    & 1  & 4              & 24 \\
\hline
9             & 9             & 36   & 9  & 36   & 9  & 36   & 9  & 36   & 9   & 36    & 9   & 36  & ---            & 24 \\
\hline
10           & 10             & 10  & 10  & 10  & 10  & 10  & 10  & 10  & 10   & 10   & 10   & 10  & ---            & 24 \\
\hline
11           & 11             & 31  & 26  & 16  & 41   & 1  & 11  & 31  & 26   & 16   & 41    & 1  & 6              & 24 \\
\hline
12           & 12              & 9  & 18  & 36  & 27   & 9  & 18  & 36  & 27    & 9   & 18   & 36  & ---            & 24 \\
\hline
\end{tabular}
\end{center}
\end{table}
} % \small

@@

{\bf Example 4:} \\
The following table\footnote{Mathematica can be used to calculate this as follows: \\
{\tt
m = 46; Do[ Print[ Table[ Mod[${\rm a}^\wedge$i, m], \{i, 1, 23\} ], \\
'', '', MultiplicativeOrder[a, m, 1], \\
'', '', EulerPhi[m] ], \\
\{a, 1, 23\} ];
} % \tt
} answers the question as to whether the number $7$ is a primitive root of $46$.
The requirement/condition $gcd(7,46)=1$ is fulfilled.
%\newpage

>>>>

Cryptographically not only the uniqueness of the inverse is important, but also that the
set of possible values is exhausted.
\begin{theorem}\label{thm-zth-exhperm}
For $a,i\in \{1, \cdots, m-1 \}$ with ${\rm~gcd} (a,m)=1),$ then takes the product $a*i {\rm ~mod ~} m $ for a 
certain number $a$ all values from $\{1, \cdots ,m-1 \} $ (exhaustive permutation \index{Permutation} of the Length $m-1$)\footnote{See
also theorem \ref{thm-zth-ordp} in \hyperlink{Kap_3_9}{chapter 3.9 Multiplicative order and primitive roots}.}.
\end{theorem}
\vskip +10 pt
The following three examples\footnote{%
See \hyperlink{AppArith1}{Appendix D} for the source code to compute the tables using Mathematica\index{Mathematica}
or Pari-GP.\index{Pari-GP}
}
illustrate the properties of multiplicative inverses.

@@

{
\begin{table}[h] \label{SrcArith1a}
\subsubsection*{Multiplication table modulo $17$ (for $a=5$ and $a=6$)}
\begin{center}
\begin{tabular}{|l||@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}||@{\:}c@{\:}|@{\:}c@{\:}|}
\hline 
i                   & 1  & 2  & 3  & 4  & 5  & 6  & 7  & 8  & 9 & 10 & 11 & 12 & 13 & 14 & 15 & 16  & 17 & 18 \\
\hline
\hline  
$5*i$                & 5 & 10 & 15 & 20 & 25 & 30 & 35 & 40 & 45 & 50 & 55 & 60 & 65 & 70 & 75 & 80  & 85 & 90   \\
remainder                & 5 & 10 & 15  & 3  & 8 & 13  & {\bf 1}  & 6 & 11 & 16  & 4  & 9 & 14  & 2  & 7 & 12 & 0  & 5   \\
\hline
$6*i$                 & 6 & 12 & 18 & 24 & 30 & 36 & 42 & 48 & 54 & 60 & 66 & 72 & 78 & 84 & 90 & 96 & 102 & 108   \\
remainder                & 6 & 12  & {\bf 1}  & 7 & 13  & 2  & 8 & 14  & 3  & 9 & 15  & 4 & 10 & 16  & 5 & 11 & 0  & 6   \\
\hline
\end{tabular}
\end{center} 
\end{table}
}
Between $i=1, \cdots, m$, all values between $0, \cdots, m-1$ occur for the remainders, because both $5$ and $6$ are 
also relatively prime\index{Relatively prime} to the modulus $m=17$.
\enlargethispage{0.5cm}

@@

\begin{table}[h] \label{SrcArith1b}
\subsubsection*{Multiplication table modulo $13$ (for $a=5$ and $a=6$)}
\begin{center}                                                                          
\begin{tabular}{|l||@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}||@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|@{\:}c@{\:}|}
\hline 
i                    & 1  & 2  & 3  & 4  & 5  & 6  & 7  & 8  & 9 & 10 & 11 & 12 & 13 & 14 & 15 & 16  & 17  & 18 \\
\hline 
\hline 
$5*i$                  & 5 & 10 & 15 & 20 & 25 & 30 & 35 & 40 & 45 & 50 & 55 & 60 & 65 & 70 & 75 & 80 & 85  & 90 \\
remainder                 & 5 & 10  & 2  & 7  & 12  & 4 & 9  & {\bf 1}  & 6  & 11 & 3  & 8  & 0 & 5  & 10  & 2   & 7   & 12 \\
\hline 
$6*i$                  & 6 & 12 & 18 & 24 & 30 & 36 & 42 & 48 & 54 & 60 & 66 & 72 & 78 & 84 & 90 & 96 & 102 & 108 \\
remainder                 & 6  & 12  & 5  & 11  & 4  & 10  & 3  & 9  & 2  & 8  & {\bf 1}  & 7  & 0  & 6  & 12  & 5   & 11   & 4 \\
\hline 
\end{tabular}
\end{center} 
\end{table}
}

@@

and the squares (mod $103$) can be calculated beforehand
\begin{eqnarray*}
87^2 & \equiv & 50 {\rm ~(mod~}103),\\
87^4 & \equiv & 50^2 \equiv 28 {\rm ~(mod~}103), \\
87^8 & \equiv & 28^2 \equiv 63 {\rm ~(mod~}103), \\
87^{16} & \equiv & 63^2 \equiv 55 {\rm ~(mod~}103),\\
87^{32} & \equiv & 55^2 \equiv 38 {\rm ~(mod~}103).
\end{eqnarray*}
we have\footnote{%
  See \hyperlink{AppArith2}{Appendix D} for source code implementing the
  square and multiply method in Mathematica and Pari-GP, which can be used
  to reproduce the calculations above. }:
\label{SrcArith2}
\begin{eqnarray}
87^{43} & \equiv & 87^{32+8+2+1}{\rm ~(mod~}103) \nonumber \\
        & \equiv & 87^{32} * 87^8 * 87^2 * 87 {\rm ~(mod~}103) \nonumber \\ 
    & \equiv & 38 * 63 * 50 * 87 \equiv 85 {\rm ~(mod~}103). \nonumber
\end{eqnarray}

@@

{\bf Example 1:} \\
The following table\footnote{%
See \hyperlink{AppArith3a}{Appendix D} for the source code to generate the table using Mathematica or Pari-GP.
} 
shows the values $a^i$ mod $11$ for the exponents $i = 1, 2, \cdots, 10$ and for the bases 
$a =  1, 2, \cdots, 10$ as well as the resulting value $ord_{11}(a)$ for each $a$:

{
\begin{table}[h]  \label{SrcArith3a}
\subsubsection*{Values of $a^i {\rm ~mod~} 11,  1 \leq a,i<11$ and according order of $a$ mod $m$:}
\begin{center}                                                                          
\begin{tabular}{|l||c|c|c|c|c|c|c|c|c|c|c|c|c|c|}
\hline
              & i=1 & i=2 & i=3 & i=4 & i=5 & i=6 & i=7 & i=8 & i=9 & i=10  & $ord_{11}(a)$\\
\hline
\hline
$a=1$           & 1  & 1    & 1  & 1    & 1    & 1    & 1  & 1    & 1  & 1     & 1   \\
\hline
$a=2$           & 2  & 4    & 8  & 5   & 10    & 9    & 7  & 3    & 6  & 1    & 10  \\
\hline
$a=3$           & 3  & 9    & 5  & 4 & {\bf 1} & 3    & 9  & 5    & 4  & 1     & 5   \\
\hline
$a=4$           & 4  & 5    & 9  & 3 & {\bf 1} & 4    & 5  & 9    & 3  & 1    & 5 \\
\hline
$a=5$           & 5  & 3    & 4  & 9 & {\bf 1} & 5    & 3  & 4    & 9  & 1    & 5   \\
\hline
$a=6$           & 6  & 3    & 7  & 9   & 10    & 5    & 8  & 4    & 2  & 1    & 10  \\
\hline
$a=7$           & 7  & 5    & 2  & 3   & 10    & 4    & 6  & 9    & 8  & 1    & 10  \\
\hline
$a=8$           & 8  & 9    & 6  & 4   & 10    & 3    & 2  & 5    & 7  & 1    & 10  \\
\hline
$a=9$           & 9  & 4    & 3  & 5 & {\bf 1} & 9    & 4  & 3    & 5  & 1    & 5   \\
\hline
$a=10$         & 10  & 1   & 10  & 1   & 10    & 1   & 10  & 1   & 10  & 1    & 2   \\
\hline
\end{tabular}
\end{center}
\end{table}
}

@@

{\bf Example 2:} \\
The following table\footnote{%
See \hyperlink{AppArith3b}{Appendix D} for the source code to generate the table using Mathematica or Pari-GP.
} shows the values $a^i$ mod $45$ for the exponents $i = 1, 2, \cdots, 12$ and for the bases 
$a =  1, 2, \cdots, 12$ as well as the resulting value $ord_{45}(a)$ for each $a$:

{ % \small
\begin{table}[h] \label{SrcArith3b}
\subsubsection*{Values of $a^i{\rm ~mod~}45, 1\leq a,i<13$:}
\begin{center}                                                                          
\begin{tabular}{|l||c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}
\hline
 $a\setminus i$ & 1            & 2            & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12     & $ord_{45}(a)$       & $J(45)$ \\
\hline
\hline                                                       
1             & 1              & 1   & 1   & 1   & 1   & 1   & 1   & 1   & 1    & 1    & 1    & 1 & 1              & 24  \\
\hline
2             & 2              & 4   & 8  & 16  & 32  & 19  & 38  & 31  & 17   & 34   & 23    & 1 & 12             & 24 \\
\hline
3             & 3              & 9  & 27  & 36  & 18   & 9  & 27  & 36  & 18    & 9   & 27   & 36  & ---            & 24 \\
\hline
4             & 4             & 16  & 19  & 31  & 34   & 1   & 4  & 16  & 19   & 31   & 34    & 1  & 6              & 24 \\
\hline
5             & 5             & 25  & 35  & 40  & 20  & 10   & 5  & 25  & 35   & 40   & 20   & 10  & ---            & 24 \\
\hline
6             & 6             & 36  & 36  & 36  & 36  & 36  & 36  & 36  & 36   & 36   & 36   & 36  & ---            & 24 \\
\hline
7             & 7              & 4  & 28  & 16  & 22  & 19  & 43  & 31  & 37   & 34   & 13    & 1  & 12             & 24 \\
\hline
8             & 8             & 19  & 17   & 1   & 8  & 19  & 17   & 1   & 8   & 19   & 17    & 1  & 4              & 24 \\
\hline
9             & 9             & 36   & 9  & 36   & 9  & 36   & 9  & 36   & 9   & 36    & 9   & 36  & ---            & 24 \\
\hline
10           & 10             & 10  & 10  & 10  & 10  & 10  & 10  & 10  & 10   & 10   & 10   & 10  & ---            & 24 \\
\hline
11           & 11             & 31  & 26  & 16  & 41   & 1  & 11  & 31  & 26   & 16   & 41    & 1  & 6              & 24 \\
\hline
12           & 12              & 9  & 18  & 36  & 27   & 9  & 18  & 36  & 27    & 9   & 18   & 36  & ---            & 24 \\
\hline
\end{tabular}
\end{center}
\end{table}
} % \small

@@

{\bf Example 4:} \\
The following table\footnote{%
See \hyperlink{AppArith3c}{Appendix D} for the source code to generate the table using Mathematica
or Pari-GP.
} answers the question as to whether the number $7$ is a primitive root of $46$.
The requirement/condition $gcd(7,46)=1$ is fulfilled.
%\newpage

\subsubsection*{Values of $a^i{\rm ~mod~}46, 1\leq a,i<23$:}
{ \textmd \small
\begin{center}  \label{SrcArith3c}
\begin{tabular}{|p{16 pt}||@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|c|}
\hline
$a \setminus i$   & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 & 13 & 14 & 15 & 16 & 17 & 18 & 19 & 20 & 21 & 22 & 23 & ord \\
\hline
\hline
1    & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1 & 1    \\
\hline
2 & 2  & 4  & 8 & 16 & 32 & 18 & 36 & 26  & 6 & 12 & 24  & 2  & 4  & 8 & 16 & 32 & 18 & 36 & 26  & 6 & 12 & 24  & 2 & --    \\
\hline
3 & 3  & 9 & 27 & 35 & 13 & 39 & 25 & 29 & 41 & 31  & 1  & 3  & 9 & 27 & 35 & 13 & 39 & 25 & 29 & 41 & 31  & 1  & 3 & 11   \\
\hline
4  & 4 & 16 & 18 & 26 & 12  & 2  & 8 & 32 & 36  & 6 & 24  & 4 & 16 & 18 & 26 & 12  & 2  & 8 & 32 & 36  & 6 & 24  & 4 & --  \\
\hline
5 & 5 & 25 & 33 & 27 & 43 & 31 & 17 & 39 & 11  & 9 & 45 & 41 & 21 & 13 & 19  & 3 & 15 & 29  & 7 & 35 & 37  & 1  & 5 & 22  \\
\hline
6 & 6 & 36 & 32  & 8  & 2 & 12 & 26 & 18 & 16  & 4 & 24  & 6 & 36 & 32  & 8  & 2 & 12 & 26 & 18 & 16  & 4 & 24  & 6 & -- \\
\hline
7 & 7  & 3 & 21  & 9 & 17 & 27  & 5 & 35 & 15 & 13 & 45 & 39 & 43 & 25 & 37 & 29 & 19 & 41 & 11 & 31 & 33  & 1  & 7 & 22 \\
\hline
8 & 8 & 18  & 6  & 2 & 16 & 36 & 12  & 4 & 32 & 26 & 24  & 8 & 18  & 6  & 2 & 16 & 36 & 12  & 4 & 32 & 26 & 24  & 8 & --  \\
\hline
9 & 9 & 35 & 39 & 29 & 31  & 3 & 27 & 13 & 25 & 41  & 1  & 9 & 35 & 39 & 29 & 31  & 3 & 27 & 13 & 25 & 41  & 1  & 9 & 11  \\
\hline
10 & 10  & 8 & 34 & 18 & 42  & 6 & 14  & 2 & 20 & 16 & 22 & 36 & 38 & 12 & 28  & 4 & 40 & 32 & 44 & 26 & 30 & 24 & 10 & --  \\
\hline 
11 & 11 & 29 & 43 & 13  & 5  & 9  & 7 & 31 & 19 & 25 & 45 & 35 & 17  & 3 & 33 & 41 & 37 & 39 & 15 & 27 & 21  & 1 & 11 & 22 \\
\hline
12 & 12  & 6 & 26 & 36 & 18 & 32 & 16  & 8  & 4  & 2 & 24 & 12  & 6 & 26 & 36 & 18 & 32 & 16  & 8  & 4  & 2 & 24 & 12 & -- \\
\hline
13 & 13 & 31 & 35 & 41 & 27 & 29  & 9 & 25  & 3 & 39  & 1 & 13 & 31 & 35 & 41 & 27 & 29  & 9 & 25  & 3 & 39  & 1 & 13 & 11  \\
\hline
14 & 14 & 12 & 30  & 6 & 38 & 26 & 42 & 36 & 44 & 18 & 22 & 32 & 34 & 16 & 40  & 8 & 20  & 4 & 10  & 2 & 28 & 24 & 14 & -- \\
\hline
15 & 15 & 41 & 17 & 25  & 7 & 13 & 11 & 27 & 37  & 3 & 45 & 31  & 5 & 29 & 21 & 39 & 33 & 35 & 19  & 9 & 43  & 1 & 15 & 22 \\
\hline
16 & 16 & 26  & 2 & 32  & 6  & 4 & 18 & 12  & 8 & 36 & 24 & 16 & 26  & 2 & 32  & 6  & 4 & 18 & 12  & 8 & 36 & 24 & 16 & -- \\
\hline
17 & 17 & 13 & 37 & 31 & 21 & 35 & 43 & 41  & 7 & 27 & 45 & 29 & 33  & 9 & 15 & 25 & 11  & 3  & 5 & 39 & 19  & 1 & 17 & 22 \\
\hline
18 & 18  & 2 & 36  & 4 & 26  & 8  & 6 & 16 & 12 & 32 & 24 & 18  & 2 & 36  & 4 & 26  & 8  & 6 & 16 & 12 & 32 & 24 & 18 & -- \\
\hline
19  & 19 & 39  & 5  & 3 & 11 & 25 & 15  & 9 & 33 & 29 & 45 & 27  & 7 & 41 & 43 & 35 & 21 & 31 & 37 & 13 & 17  & 1 & 19 & 22 \\
\hline
20  & 20 & 32 & 42 & 12 & 10 & 16 & 44  & 6 & 28  & 8 & 22 & 26 & 14  & 4 & 34 & 36 & 30  & 2 & 40 & 18 & 38 & 24 & 20 & --  \\
\hline
21 & 21 & 27 & 15 & 39 & 37 & 41 & 33  & 3 & 17 & 35 & 45 & 25 & 19 & 31  & 7  & 9  & 5 & 13 & 43 & 29 & 11  & 1 & 21 & 22  \\
\hline
22 & 22 & 24 & 22 & 24 & 22 & 24 & 22 & 24 & 22 & 24 & 22 & 24 & 22 & 24 & 22 & 24 & 22 & 24 & 22 & 24 & 22 & 24 & 22 & -- \\
\hline
23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & -- \\
\hline 
\end{tabular}
\end{center}
}
